---
title: "Using BactDating to test the clock model"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using BactDating to test the clock model}
  %\usepackage[utf8]{inputenc}
---

##Initialisation

```{r}
library(BactDating)
library(ape)
set.seed(0)
```

##Data generated from strict clock model

We start by generating a coalescent tree with 10 leaves sampled at regular intervals between 1990 and 2010, and a coalescent time unit of 5 years:

```{r}
dates=1990:2010
phy=simcoaltree(dates,alpha=5)
plot(phy,show.tip.label = F)
axisPhylo(backward = F)
```

On each branch we observe a number of substitutions which is distributed $\mathrm{Gamma}(rl,1)$
where $l$ is the branch length and 
$r=10$ per year is the substitution rate. We can simulate an observed phylogenetic tree
and perform a root-to-tip analysis as follows:

```{r}
obsphy=simobsphy(phy,mu=10)
res=roottotip(obsphy,dates)
```
##Analysis of data from strict clock model

We run the mixture-model dating analysis as follows:

```{r}
res=bactdate(obsphy,dates,updateRoot=F)
plot(res,'trace')
print(res$pstrict)
```

##Data generated from relaxed clock model

Let's start again with a new dataset using the relaxed model:
```{r}
set.seed(0)
obsphy=simobsphy(phy,mu=10,model='relaxedgamma',sigma = 5)
res=roottotip(obsphy,dates)
```

#Analysis of data from relaxed clock model

We run the analysis as previously:

```{r}
res=bactdate(obsphy,dates,updateRoot=F)
plot(res,'trace')
print(res$pstrict)
```
