---
title: "Using BactDating after recombination analysis"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Using BactDating after recombination analysis}
  %\usepackage[utf8]{inputenc}
---

## Initialisation

```{r}
library(BactDating)
library(ape)
set.seed(0)
```

## Data

A ClonalFrameML output can be read using:

```{r eval=FALSE}
t=loadCFML(prefix='cfml-output')
```

A Gubbins output can be read using:

```{r eval=FALSE}
t=loadGubbins(prefix='gubbins-output')
```


Instead, for the purpose of making this vignette self-contained, we will simulate an output from a recombination analysis.
This is just an object of class `phylo`, unrooted and with the added field `unrec`
representing the proportion that is unrecombined for each branch.
```{r}
dates=1990:2010
truetree=simcoaltree(dates)
t=unroot(simobsphy(truetree))
t$unrec=runif(length(t$edge.length))
plot(t)
axisPhylo(backward = F)
edgelabels(round(100*t$unrec))
```

## Analysis

We run BactDating as follows:

```{r}
res=bactdate(t,dates,useRec=T)
plot(res,'treeCI')
```

We can compare with the correct tree:
```{r}
plot(truetree)
axisPhylo(backward = F)
```

We can see what the MCMC traces look like:

```{r}
plot(res,'trace')
```
