---
title: "Example of using CreDating"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

##Initialisation

```{r}
library(CreDating)
library(ape)
set.seed(0)
```

##Data

We start by generating a random timed tree with 20 leaves and root in 2000:

```{r}
phy <- simdatedtree(nsam = 20, dateroot = 2000)
plot(phy)
axisPhylo(backward = F)
```

We can extract the sampling dates as follows:
```{r}
dates = leafDates(phy)
```

On each branch we observe a number of substitutions which is distributed $\mathrm{Gamma}(rl,1)$
where $l$ is the branch length and 
$r$ is the substitution rate specific to this branch, which has mean 10 substitutions per year. 
We can simulate an observed phylogenetic tree
and perform a root-to-tip analysis as follows:
```{r}
obsphy=simobsphy(phy, rate=10,model='relaxedgamma',ratevar = 100)
res=roottotip(obsphy,dates,predInt='gamma',showTree=T)
```

# ##Unknown root
#Let's say we don't know where the root is, we might consider as a guess that it is at
#the midpoint:
#```{r}
#obsphy=phangorn::midpoint(obsphy)
#res=roottotip(obsphy, dates, predInt = 'gamma', showTree = T)
#```

##Analysis

We run CreDating as follows:

```{r}
res=credate(obsphy,dates,nbIts = 10000,model='relaxedgamma',updateRoot = 2,showProgress=F,initRate=10,initRatevar = 100,updateRatevar = T)
plot(res)
```

Plot with credibility bars:
```{r}
plot(res,'treeCI')
```


We can see what the MCMC traces look like:

```{r}
plot(res,'trace')
```
Let's see where the root is likely to be:
```{r}
plot(res,'treeRoot')
```
