---
title: "Example of using CreDating with relaxed clock model"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
---

##Initialisation

```{r}
library(CreDating)
library(ape)
set.seed(0)
```

##Data generated from strict clock model

We start by generating a coalescent tree with 10 leaves sampled in 2000, 10 leaves sampled in 2010, and a coalescent time unit of 10 years:

```{r}
dates=c(rep(2000,10),rep(2010,10))
phy=simcoaltree(dates,alpha=10)
plot(phy,show.tip.label = F)
axisPhylo(backward = F)
```

On each branch we observe a number of substitutions which is distributed $\mathrm{Gamma}(rl,1)$
where $l$ is the branch length and 
$r=10$ per year is the substitution rate. We can simulate an observed phylogenetic tree
and perform a root-to-tip analysis as follows:

```{r}
obsphy=simobsphy(phy,rate=10)
res=roottotip(obsphy,dates)
```
##Analysis of data from strict clock model

We run the mixture-model dating analysis as follows:

```{r}
res=credate(obsphy,dates,model='mixedgamma',updateRoot=F)
plot(res,'trace')
print(res$pstrict)
```

##Data generated from relaxed clock model

Let's start again with a new dataset using the relaxed model:
```{r}
rm(list=ls())
set.seed(0)
dates=c(rep(2000,10),rep(2010,10))
phy <- simcoaltree(dates,alpha=10)
obsphy=simobsphy(phy,rate=10,model='relaxedgamma',ratevar = 100)
res=roottotip(obsphy,dates)
```

#Analysis of data from relaxed clock model

We run the analysis as previously:

```{r}
res=credate(obsphy,dates,model='mixedgamma',updateRoot=F)
plot(res,'trace')
print(res$pstrict)
```
