---
title: "Using BactDating on a real dataset"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Application of BactDating on S. aureus ST239}
  %\usepackage[utf8]{inputenc}
---

In this vignette we demonstrate the usage of BactDating on a dataset of
Staphylococcus aureus ST239 published by Harris et al (2010).
The data for this example can be loaded using the command `data(staph)`.
This will load in your environment the object `staph` which is made of two parts. 
The first part, `staph$tree`, is a phylogenetic tree class `phylo` from the `ape` package.
The second part, `staph$dates`, is a vector containing the dates of isolations of the genomes.
Note that this vector is in the same order as the tips of the tree listed in `staph$tree$tip.label`.

## Initialisation

```{r}
library(BactDating)
library(ape)
set.seed(0)
data(staph)
plot(staph$tree, show.tip.label = F)
axisPhylo(backward = F)
```

## Root-to-tip analysis

Let's do a root-to-tip regression analysis:

```{r}
res=roottotip(staph$tree,staph$dates)
```

## Analysis

We can run BactDating as follows:

```{r}
res=bactdate(unroot(staph$tree),staph$dates)
plot(res,'treeCI',show.tip.label = F)
```

We can see what the MCMC traces look like:

```{r}
  plot(res,'trace')
```

Let's see where the root is likely to be:

```{r}
plot(res,'treeRoot',show.tip.label=F)
```
