---
title: "Using BactDating on your own data"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Application of BactDating on your own data}
  %\usepackage[utf8]{inputenc}
---

In this vignette we describe how you can use BactDating on your own data.

## First input: phylogenetic tree

The first required input (let's call it `t`) is a phylogenetic tree
of the class `phylo` from the `ape` package,
which can be loaded from a Newick or Nexus file using respectively the following commands
from the `ape` package: 

```{r eval=FALSE}
t=read.tree('filename.nwk')
t=read.nexus('filename.nex')
```

This tree should have branch lengths measured in units of substitutions.
In other words, the branch lengths should not be measured in substitutions per site. 
If you are not sure whether the branch lengths are measured correctly or not, try
to compute `sum(t$edge.length)`. This should be the total number of substitutions throughout the tree,
so if you get a value below one or even only a bit above one, your branch lengths are probably
not in the correct unit.

If the branch lengths are measured per site, you will need to rescale them using the command:
```{r eval=F}
t$edge.length=t$edge.length*L
```
where `L` is the number of sites which were used to build this tree. 

Alternatively, you can load directly from the output of ClonalFrameML or Gubbins using
the following commands respectively:
```{r eval=F}
t=loadCFML(prefix)
t=loadGubbins(prefix)
```
This will have the advantage to account for the recombination events detected by ClonalFrameML
or Gubbins when performing the dating.

## Second input: isolation dates

The second required input (let's call it `d`) is the dates at which the isolates were sampled. 
Dates need to be expressed decimally, for example `2015.5` means 1st July 2015.
This input can be simply a vector of decimal dates, in which case the order needs to be the same
as the order of the tree tips in `t$tip.label`. For example, if `t$tip.label` is `(A,B,C)` 
then `d=c(2010,2011,2012)` indicates that `A` was isolated in 2010, B in 2011 and C in 2012.

Alternatively, the vector `d` does not have to be ordered if `names(d)` corresponds to the tip names.
For example, d=c(2011,2010,2012)` with names(d)=c('B','A','C')` would have the same meaning
as in the previous example.

If some isolation dates are unknown, this can be indicated using `NA` in the vector, in 
which case the date is considered unknown but within the range of known dates,
ie with the range return by the command `range(d,na.rm=T)`.

Finally, the isolation dates can be specified as an interval rather than a set value. This is useful
for example if the isolation year is 2015 but with unknown month or day. To do so, `d` needs
to be a matrix with two columns, with the first column indicating the lower bound and the second 
column indicating the upper bound of the interval. 

## Running commands

Once the two inputs `t` and `d` described above are in the correct format, 
you are ready to run BactDating using the main command:
```{r,eval=F}
res=bactdate(t,d)
```
See other vignettes for further functionality. 

## Checking for MCMC convergence

BactDating is a MCMC algorithm which needs to be run for long enough to ensure that the 
results are meaningful. The length of the algorithm can be controlled using the `nbIts` argument
of the `bactdate` function, which by default is 10e4 but may need increasing for example ten folds:
```{r,eval=F}
res=bactdate(t,d,nbIts=1e5)
```

To test the MCMC convergence, a good start is to check that the traces of the parameters look stable 
using the command:
```{r,eval=F}
plot(res,'trace')
```

For further testing of convergence, you can export the BactDating result to the format required
by the `coda` package using the command:
```{r,eval=F}
mcmc=as.mcmc(res)
```

You can then compute for example the effective sample size of the parameters which should be at least 100 for each paramater, using:
```{r,eval=F}
effectiveSize(mcmc)
```
